---
title: "omics_exs"
author: "Marta Españó López"
date: "5/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## R Markdown

EXERCISE 2 OMICS


Link to my github repository:
https://github.com/Martaesplo/omics_techniques.git
It contains the html file too.


- Study: 	
Ethanol treatment of lymphoblastoid cell lines from alcoholics and non-alcoholics causes many subtle changes in gene expression

- Accession number: 
GSE52553

- Description of the study: 
To understand the effect of controlled exposure to ethanol on gene expression, lymphoblastoid cell lines (LCLs) from 21 alcoholics and 21 controls were studied after a treatment of 24 hours. Gene expression was meassured using microarrays which together with a paired design allowed to detect very small changes between the cell lines due to ethanol treatment.

- Information of the experimental design:
As mentioned before is paired design. It is composed of two factors and two groups: treatment/no-treatment and alcoholic/non-alcoholic(control). 
The sample size is 84 and has biological origin.
In the study the variable sex (female or male) is also specified, but we will ignore it.
The response variable is gene expression and it depends on the ethanol treatment and if the sample is from an alcoholic LCL or not.
And is a block design as the experimental units are in groups (blocks).


Installation of GEOquery:

```{r geoquery}
library(methods)
library(Biobase)
library(BiocManager)
library(BiocVersion)
  if (!require(GEOquery)) {
  BiocManager::install("GEOquery")
}
BiocManager::install("GEOquery")
require(GEOquery)
gse <- getGEO("GSE52553")
```


- Brief data description:

```{r data_description}
class(gse)
names(gse)
gse[[1]]
expg<-gse[[1]]
dim(exprs(expg))
class(phenoData(expg))
head(pData(expg))

targets <- data.frame(gse)
class(targets)
length(targets)
names(targets)
head(targets) #all the samples
m <- matrix(rnorm(targets))
hist(m, main='Histogram targets', xlab='targets', col='purple')
boxplot(m, main='Boxplot targets', col='purple')

```
