---
title: "synthesis exercise"
author: "Marta Españó López"
date: "6/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Marta Espano Lopez
104532


Link to my github repository:
https://github.com/Martaesplo/omics_techniques.git


GEO accession number: GSE52553


Installation and libraries needed:

```{r packages, echo=FALSE}
installifnot <- function (pkg){
  if (!require(pkg, character.only=T)){
  source("http://bioconductor.org/install")
  biocLite(pkg)
}else{
  require(pkg, character.only=T)
  }
}

installifnot("pd.mogene.1.0.st.v1")
BiocManager::install("mogene10sttranscriptcluster.db")
installifnot("oligo")
installifnot("limma")
installifnot("Biobase")
# installifnot("arrayQualityMetrics")
BiocManager::install("genefilter")
BiocManager::install("multtest")
BiocManager::install("annotate")
#install("gplots")
installifnot("scatterplot3d")
```

```{r libraries, echo=FALSE}
library(methods)
library(Biobase)
library(BiocManager)
library(BiocVersion)
library(GEOquery)
library(limma)
library(pd.mogene.1.0.st.v1)
library(mogene10sttranscriptcluster.db)
library(oligo)
library(limma)
library(genefilter)
library(multtest)
library(annotate)
library(gplots)
library(scatterplot3d)
```


DATA DESCRIPTION

- Study: 	
Ethanol treatment of lymphoblastoid cell lines from alcoholics and non-alcoholics causes many subtle changes in gene expression.

- Description of the study: 
To understand the effect of controlled exposure to ethanol on gene expression, lymphoblastoid cell lines (LCLs) from 21 alcoholics and 21 controls were studied after a treatment of 24 hours. Gene expression was meassured using microarrays which together with a paired design allowed to detect very small changes between the cell lines due to ethanol treatment.


First we download the GEO dataset GSE52553 and since is has a sample size of 84, we choose a subset of it containing 12 samples. Those are balanced according to the existent groups.


```{r description, echo=FALSE}
gse <- getGEO("GSE52553")
eset<-gse[[1]]
rdeset<-(eset[1:11,c(1,2,3,4,43,44,65,66,69,70,29,30)])

dim(exprs(rdeset))
dim(exprs(eset))
head(rdeset)
exprs(rdeset)
head(exprs(rdeset))
targets<-pData(rdeset)
targets
```


```{r plots, echo=TRUE}

print(paste0('Mean of st.deviation = ',mean(pcs$sdev)))

pcs<-prcomp(exprs(rdeset))
barplot(pcs$sdev, col='purple', main='Un-normalized expression set')

plot(pcs$rotation[,1], pcs$rotation[,2], main="Representation of first two principal components", col='purple')

boxplot(pcs, col='purple', main='Boxplot expression set')

```



- Information of the experimental design:

Response variable: Gene expression
Depends on: ethanol Treatment and being alcoholic or not
Factors and groups: Treatment/no treatment and alcoholic/non-alcoholic(control)
Sample size: 12 (84 originally), 3 samples belonging to each group
Experimental units per combination of factors: 3
Type of experimental design: Randomized block design
Replicate type: Biological

In the study the variable sex (female or male) is also specified, but we will ignore it.


-Comparisons to be performed:
Between the parenthesis there are the number of samples of each group.
Since we want to know the effect of both, alcohol and treatment with ethanol, to gene expression we will have to perform the following comparisons.










